Alleviating the Topology Mismatch Problem in Distributed Overlay Networks: A Survey

### 2015-09-26 structured.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

Even though \emph{traceroute}
provides detailed information about the network structure, 
its extensive use may create overhead on the overall network structure. 
Administrators thus frequently
disable \emph{traceroute} support on their routers and this 
undoubtly affects \emph{T2MC}'s operation.
%%
Although overall experimental results have been encouraging, 
\emph{T2MC} is considered rather coarse-grained as it only
classifies neighbors into two groups \cite{QLZG2009}.
%
%T2MC \cite{SLCGZ2008} uses trace-route logs to detect different autonomous
%systems and cluster close by peers, thus reducing redundant traffic that crosses
%autonomous system boundaries. T2MC uses a customized k-mean classification
%algorithm with $k=2$ to perform the classification, and exploits the stable
%structure of the Internet routers to guide clustering. The peer chooses the
%minimum and maximum Latency results from the trace-route to initialize the
%centroids of two sets. Then the peer calculates, for all hops along its
%trace-routed path, the peer calculates the absolute distance to these centroids
%of both sets and assigns the routers to that centroid with which it has the
%smaller absolute distance. After that, the mean and variance values are
%calculated and if the variance is larger than a predefined threshold then the
%algorithm sets the two latency mean values as new centroids for the two sets and
%loops again. Ultimately peer will end up with two sets having the minimum
%intra-set variance. The peer, then, chooses the router from the second set with
%the minimum hops attribute and sets it as a threshold. The selected router and
%all others whose hops attribute is larger than the threshold are classified as
%\emph{remote} router cluster. The remaining are classified as \emph{near}. From
%the near class, the peer chooses the one with maximum hops attribute as its edge
%router, and registers it along with all near clusters into the DHT of the
%\p\ overlay. As new peers join the network, those that share the same edge
%router or any of the members of the near router clusters, would form a close
%peer cluster. As Edge routers can provide more valuable information than other
%members of the near set, T2MC was designed to prioritize interaction of peers
%and edge gateways.
%
%
% TODO: SOME DISCUSSION
%
%The use of Trace-route as a tool for implementing the distance measuring
%infrastructure raise concerns about its efficiency and scalability. Being,
%primarily, a network diagnostic utility,
% Even though trace-route provides detailed information about the network
% structure, its excessive use creates additional overhead that stresses further
% the overall network infrastructure. This forces network administrators to
% disable its support from the routers they manage, thus jeopardizing the
% effectiveness of the T2MC algorithm.
%
\emph{T2MC} fares as follows:
%%
\begin{center}
{\footnotesize
\begin{tabular}{rccc}
\multicolumn{1}{r}{} &
\multicolumn{1}{c}{\emph{Efficiency}} &
\multicolumn{1}{c}{\emph{Overhead}} &
\multicolumn{1}{c}{\emph{Scalability}}
\\
\cline{2-4}
\emph{T2MC} &
% The efficiency of choosing close by peers is very good and reported almost
% 65\% better than that of GNP (and thus generally of static landmarking as a
% whole)
high &
% trace-route is a very heavy tool and thus incurs excessive burden to the
% network resources.
high &
% Trace-route is categorized as too heavy-weighted and intrusive for use in a
% larger scale system \cite{RKHS2002}. Additionally, disabling
% ICMP is a common administrative policy for edge sites to enforce security,
% while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not
% directly available to the application layer.
low
\end{tabular}
}
\end{center}

@@

\emph{CNPA} is proposed for network setups for which a lightweight
bootstrap phase is required.
It delegates the task of identifying the physically
close--node in a \emph{Pastry}-based network; this node is used by the 
incoming peer to get bootstrapped.
The new-comer takes the bootstrapping nodes's \emph{ID} prefix 
and fills the remaining bits of its key using the \emph{RLM} technique.
%%
\emph{RLM} is evaluated and exhibits good proximity in load-balanced 
environments. Moreover, it is adaptable to high-churn settings.
Unfortunately, \emph{RLM} bootstrap overhead is substantial and can 
effect the set of landmarks~\cite{WZS2004}.
\emph{CNPA} on the other hand, incurs less overhead but this comes 
at the expense of being more coarse-grained as it is based on prefixes
to determine proximity during bootstrap.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \emph{RLM} achieves good proximity evaluation in a load balanced and adaptable to
% high churn environments fashion. Unfortunately it comes with more bootstrap
% overhead that can especially effect the set of temporary landmarks. On the other
% hand, \emph{CNPA}'s less incurred overhead comes at the expense of being more
% coarse-grained.
%%%
%%
% \emph{DynaMO} tries to capitalize on the fact that in Pastry each node's table consists
% of rows equal to the number of digits in the overlay ID and columns equal to
% the ID's base. As we go down the table rows, the matching prefix between the
% current node's ID and the row's entries increases by one. Thus, the leaf set
% contains the closest nodes in the ID space. Additionally, as the prefix match
% increases by one, the result is exponentially fewer candidates that can fill the
% tables entries as the row increases. This leads to the observation
%  In this context, two approaches are considered namely
% \emph{Random Landmarking (RLM)} and \emph{Closest Neighbor Prefix Assignment
% (CNPA)}.
%
% TODO: SOME DISCUSSION
%
%Both algorithms, are protected against the formation of physical landmark
%  clusters or imbalanced ID distribution\footnote{More common during the
% initial formation of an overlay network} , by introducing the \emph{landmark
% gravitation range} as a threshold over which landmark keys are reassigned (for
% the RLM approach) or unutilized ID prefix ranges are detected and used (for
% the CPNA scheme) in order to balance the distribution of regions in the
% overlay.
%
\emph{DynoMO}'s elements fare as follows regarding our three criteria:
\begin{center}
{\footnotesize
\begin{tabular}{rccc}
\multicolumn{1}{r}{} &
\multicolumn{1}{c}{\emph{Efficiency}} &
\multicolumn{1}{c}{\emph{Overhead}} &
\multicolumn{1}{c}{\emph{Scalability}}
\\
\cline{2-4}
\emph{RLM} &
%%
%% TAKEN FROM
%% Efficient service discovery mechanism for wireless sensor networks
%%
%%
% common prefix of the overlay ID with the closest landmark. The
% RLM utilizes the overlay lookup capabilities by reducing traffic
% overhead within an overlay cluster. However, it generates much
% control traffic for periodically measuring distances to all landmarks.
% Moreover, the hop stretch overhead still increases when
% performing overlay routing among overlay clusters.
medium &
%
medium &
% Dynamic landmarking approach helps the scalability of the algorithm. The
% authors claim that the temporary nature of RLM's landmark nodes does not
% prevent them to maintain a good ratio in the presence of high landmark
% digression rates.
high \\
\emph{CNPA} &
%
medium &
% % The paper suggests that this approach generates significantly lower overhead
% than the vanilla Pastry algorithm
low &
% The queries are directed through fast nodes so it is important such nodes
% be part of the network
medium \\
\end{tabular}
}
\end{center}

@@

In this section, we have discussed research efforts 
trying to address the issue of inefficient network resource utilization 
manifested by network--topology mismatches in \emph{structured} \p-systems.
Based on the salient feature of outlined techniques
\cite{CDHR2002,CDCR2002,RSS2002}, we classify furnished solutions 
based on the following aspects:
\begin{enumerate}[\itshape i\upshape)]
  \item geographic-layout,
  \item proximity routing, and 
  \item proximity neighbor selection,
\end{enumerate}
%%
Trends for structured systems depicted in Figure~\ref{structured:timeline}
show the same lack of paradigm shift we identified for the unstructured
approaches (check Figure~\ref{unstructured:timeline}). Leveraging proximity
information for selecting the next hop or to construct and maintain
routing tables as a whole, seem to be the elements more used in the
literature throught the years that this area is researched.
%%%
Similarly to unstructured approaches, most of the methods
reviewed in this section entail features from multiple of the above categories. 
For example, \cite{KLKP2008}~uses both proximity routing 
and proximity neighbor selection to achieve topology adaptation,
while \emph{LAPTOP} combines geographic--layout with
replication strategies to refine its results. What looks a bit different
compared the trends in unstructred systems is the fact that, here, researchers
seem to juggle with more methods at the same time. This, and the
fact that these suggested solutions are also heuristic in nature, opens them up 
to even harder trade-offs which, in most cases, are 
highly affected by the application domain and the implementation itself.

In what follows, we place the surveyed \emph{structured} \p-methods in 
the three above categories formulated based on 
utilized proximity and geographic aspects.
Table~\ref{structured:table} outlines key properties for all surveyed approach 
that helps differentiate and better position competing techniques.
Clearly, we do not further elaborate topics that we have included and discussed 
at length in Section~\ref{section:unstructured}; these aspects entail 
topology adaptation, landmarking, and caching/replication.
Figures~\ref{structured:plot:efficiency}, \ref{structured:plot:overhead}
and \ref{structured:plot:scalability} help with the pictorial comparison
of the algorithms' behavior with respect to the three 
criteria we defined in subsection~\ref{background:motivation}. Finally,
we embed algorithms and characteristics on the temporal axis in
Figure~\ref{structured:timeline} to help depict emerging trends during
all these years of research.

>>>>

Even though \emph{traceroute}
provides detailed information about the network structure, 
its extensive use may create overhead on the overall network structure. 
Administrators thus frequently
disable \emph{traceroute} support on their routers and this 
undoubtedly affects \emph{T2MC}'s operation.
%%
Although overall experimental results have been encouraging, 
\emph{T2MC} is considered rather coarse-grained as it only
classifies neighbors into two groups \cite{QLZG2009}.
%
%T2MC \cite{SLCGZ2008} uses trace-route logs to detect different autonomous
%systems and cluster close by peers, thus reducing redundant traffic that crosses
%autonomous system boundaries. T2MC uses a customized k-mean classification
%algorithm with $k=2$ to perform the classification, and exploits the stable
%structure of the Internet routers to guide clustering. The peer chooses the
%minimum and maximum Latency results from the trace-route to initialize the
%centroids of two sets. Then the peer calculates, for all hops along its
%trace-routed path, the peer calculates the absolute distance to these centroids
%of both sets and assigns the routers to that centroid with which it has the
%smaller absolute distance. After that, the mean and variance values are
%calculated and if the variance is larger than a predefined threshold then the
%algorithm sets the two latency mean values as new centroids for the two sets and
%loops again. Ultimately peer will end up with two sets having the minimum
%intra-set variance. The peer, then, chooses the router from the second set with
%the minimum hops attribute and sets it as a threshold. The selected router and
%all others whose hops attribute is larger than the threshold are classified as
%\emph{remote} router cluster. The remaining are classified as \emph{near}. From
%the near class, the peer chooses the one with maximum hops attribute as its edge
%router, and registers it along with all near clusters into the DHT of the
%\p\ overlay. As new peers join the network, those that share the same edge
%router or any of the members of the near router clusters, would form a close
%peer cluster. As Edge routers can provide more valuable information than other
%members of the near set, T2MC was designed to prioritize interaction of peers
%and edge gateways.
%
%
% TODO: SOME DISCUSSION
%
%The use of Trace-route as a tool for implementing the distance measuring
%infrastructure raise concerns about its efficiency and scalability. Being,
%primarily, a network diagnostic utility,
% Even though trace-route provides detailed information about the network
% structure, its excessive use creates additional overhead that stresses further
% the overall network infrastructure. This forces network administrators to
% disable its support from the routers they manage, thus jeopardizing the
% effectiveness of the T2MC algorithm.
%
\emph{T2MC} fares as follows:
%%
\begin{center}
{\footnotesize
\begin{tabular}{rccc}
\multicolumn{1}{r}{} &
\multicolumn{1}{c}{\emph{Efficiency}} &
\multicolumn{1}{c}{\emph{Overhead}} &
\multicolumn{1}{c}{\emph{Scalability}}
\\
\cline{2-4}
\emph{T2MC} &
% The efficiency of choosing close by peers is very good and reported almost
% 65\% better than that of GNP (and thus generally of static landmarking as a
% whole)
high &
% trace-route is a very heavy tool and thus incurs excessive burden to the
% network resources.
high &
% Trace-route is categorized as too heavy-weighted and intrusive for use in a
% larger scale system \cite{RKHS2002}. Additionally, disabling
% ICMP is a common administrative policy for edge sites to enforce security,
% while dumping BGP routing tables\cite{krishnamurthy_bgpclust_2000} is not
% directly available to the application layer.
low
\end{tabular}
}
\end{center}

@@

\emph{CNPA} is proposed for network setups for which a lightweight
bootstrap phase is required.
It delegates the task of identifying the physically
close--node in a \emph{Pastry}-based network; this node is used by the 
incoming peer to get bootstrapped.
The new-comer takes the bootstrapping node's \emph{ID} prefix 
and fills the remaining bits of its key using the \emph{RLM} technique.
%%
\emph{RLM} is evaluated and exhibits good proximity in load-balanced 
environments. Moreover, it is adaptable to high-churn settings.
Unfortunately, \emph{RLM} bootstrap overhead is substantial and can 
effect the set of landmarks~\cite{WZS2004}.
\emph{CNPA} on the other hand, incurs less overhead but this comes 
at the expense of being more coarse-grained as it is based on prefixes
to determine proximity during bootstrap.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \emph{RLM} achieves good proximity evaluation in a load balanced and adaptable to
% high churn environments fashion. Unfortunately it comes with more bootstrap
% overhead that can especially effect the set of temporary landmarks. On the other
% hand, \emph{CNPA}'s less incurred overhead comes at the expense of being more
% coarse-grained.
%%%
%%
% \emph{DynaMO} tries to capitalize on the fact that in Pastry each node's table consists
% of rows equal to the number of digits in the overlay ID and columns equal to
% the ID's base. As we go down the table rows, the matching prefix between the
% current node's ID and the row's entries increases by one. Thus, the leaf set
% contains the closest nodes in the ID space. Additionally, as the prefix match
% increases by one, the result is exponentially fewer candidates that can fill the
% tables entries as the row increases. This leads to the observation
%  In this context, two approaches are considered namely
% \emph{Random Landmarking (RLM)} and \emph{Closest Neighbor Prefix Assignment
% (CNPA)}.
%
% TODO: SOME DISCUSSION
%
%Both algorithms, are protected against the formation of physical landmark
%  clusters or imbalanced ID distribution\footnote{More common during the
% initial formation of an overlay network} , by introducing the \emph{landmark
% gravitation range} as a threshold over which landmark keys are reassigned (for
% the RLM approach) or unutilized ID prefix ranges are detected and used (for
% the CPNA scheme) in order to balance the distribution of regions in the
% overlay.
%
\emph{DynoMO}'s elements fare as follows regarding our three criteria:
\begin{center}
{\footnotesize
\begin{tabular}{rccc}
\multicolumn{1}{r}{} &
\multicolumn{1}{c}{\emph{Efficiency}} &
\multicolumn{1}{c}{\emph{Overhead}} &
\multicolumn{1}{c}{\emph{Scalability}}
\\
\cline{2-4}
\emph{RLM} &
%%
%% TAKEN FROM
%% Efficient service discovery mechanism for wireless sensor networks
%%
%%
% common prefix of the overlay ID with the closest landmark. The
% RLM utilizes the overlay lookup capabilities by reducing traffic
% overhead within an overlay cluster. However, it generates much
% control traffic for periodically measuring distances to all landmarks.
% Moreover, the hop stretch overhead still increases when
% performing overlay routing among overlay clusters.
medium &
%
medium &
% Dynamic landmarking approach helps the scalability of the algorithm. The
% authors claim that the temporary nature of RLM's landmark nodes does not
% prevent them to maintain a good ratio in the presence of high landmark
% digression rates.
high \\
\emph{CNPA} &
%
medium &
% % The paper suggests that this approach generates significantly lower overhead
% than the vanilla Pastry algorithm
low &
% The queries are directed through fast nodes so it is important such nodes
% be part of the network
medium \\
\end{tabular}
}
\end{center}

@@

In this section, we have discussed research efforts 
trying to address the issue of inefficient network resource utilization 
manifested by network--topology mismatches in \emph{structured} \p-systems.
Based on the salient feature of outlined techniques
\cite{CDHR2002,CDCR2002,RSS2002}, we classify furnished solutions 
based on the following aspects:
\begin{enumerate}[\itshape i\upshape)]
  \item geographic-layout,
  \item proximity routing, and 
  \item proximity neighbor selection,
\end{enumerate}
%%
Trends for structured systems depicted in Figure~\ref{structured:timeline}
show the same lack of paradigm shift we identified for the unstructured
approaches (check Figure~\ref{unstructured:timeline}). Leveraging proximity
information for selecting the next hop or to construct and maintain
routing tables as a whole, are the two predominant elements used 
throughout the years in 
the structured systems literature.
%%%
Similarly to unstructured approaches, most of the methods
reviewed in this section entail features from multiple of the above categories. 
For example, \cite{KLKP2008}~uses both proximity routing 
and proximity neighbor selection to achieve topology adaptation,
while \emph{LAPTOP} combines geographic--layout with
replication strategies to refine its results. What looks different
compared with the trends in unstructured systems is the fact that, here, researchers
seem to juggle with more methods at the same time. This, and the
fact that these suggested solutions are also heuristic in nature, opens them up 
to trade-offs that are more difficult to manage and are highly 
affected by both the application domain and the implementation itself.

In what follows, we place the surveyed \emph{structured} \p-methods in 
the three above categories formulated based on 
utilized proximity and geographic aspects.
Table~\ref{structured:table} outlines key properties for all surveyed approaches 
that helps differentiate and better position competing techniques.
Clearly, we do not further elaborate topics that we have included and discussed 
at length in Section~\ref{section:unstructured}; these aspects entail 
topology adaptation, landmarking, and caching/replication.
Figures~\ref{structured:plot:efficiency}, \ref{structured:plot:overhead}
and \ref{structured:plot:scalability} help with the pictorial comparison
of the algorithms' behavior with respect to the three 
criteria we defined in subsection~\ref{background:motivation}. Finally,
we embed algorithms and characteristics on the temporal axis in
Figure~\ref{structured:timeline} to help depict emerging trends during
all these years of research.

### 2015-09-26 unstructured.tex
2015.09.28 (Alex, Mema & Vassilis) Rephrasing and fixing typos. Adding rebutal letter and new letter to area editor. Including revised (r1) submission. Minor file reorganization and Makefile cleanup.

As discussed above, \emph{THANCS} is fully distributed and needs only minimum knowledge
of, at most, a $2$-hop distant peers. This features renders 
the method a good candidate for large overlay networks.
Also its topology adaptation helps construct a well fit overlay
with respect to the underlying network--at least with respect to network
distances. In addition, will-cut list and distance caches (which store already
probed peers) minimize the unnecessary messages for the network maintenance.
%%
\emph{THANCS} reportedly improves brodcast performance the extent of which, though,
depends on the underlying network since \emph{THANCS} does not provide the
performance guarantee for the improvement of any given overlay topology~\cite{HLY2010}.

@@

While surveying efforts to overcome the mismatch problem
in the area of unstructured \p\ systems,
we came to identify four key methodologies 
utilized by the discussed approaches; they are:
\begin{enumerate}[\itshape i\upshape)]
  \item topology adaptation
  \item forwarding optimization
  \item caching and replication, and
  \item landmarking.
\end{enumerate}
%
A key characteristic shared by many of the presented techniques 
is that they do not strictly adhere to a single methodology.
As approaches attempt to address the challenging problem 
of topology mismatch, they resort to heuristics and so, 
they do not often furnish ``pure'' approaches that exclusively
belong to one of the $4$ aforementioned methodologies.
For instance, \emph{mOverlay} combines topology adaptation 
with landmark binning 
while \emph{Gia} weaves together $3$ methodologies:  topology adaptation, 
forwarding optimization and caching.
%%
The timeline in Figure~\ref{unstructured:timeline} further emphasises this trend
in the broader picture. All characteristics we discuss,
seem to have been identified and put into practice, early on in the literature.
In that sense, no breakthrough
or any revolutionary ideas came out lately but mostly people fine-tuned the
heuristic parameters in pursue of the goldset that performs best. That doesn't
render all that work less exciting or important since it led to better theoretical
understanding of the problem and its difficulty. Moreover, the combination of
multiple methodologies inevitably leads to interesting feature/performance trade-offs
that researchers need to tackle in real world applications: for example, 
the effectiveness of a caching/replication component can be undermined
by a continuously adapting overlay that removes important links between peers.
Topology adaptation and forwarding optimization seem to be constant areas of
concentration for the researchers. This is of no surprise as all seem to
understand the importance of the volatile nature of unstructured \p systems. 
%%
If a technique is to be evaluated for a specific application domain,
its overall approach as well as its underlying used methodologies have to be considered.

>>>>

As discussed above, \emph{THANCS} is fully distributed and needs only minimum knowledge
of, at most, a $2$-hop distant peers. This features renders 
the method a good candidate for large overlay networks.
Also its topology adaptation helps construct a well fit overlay
with respect to the underlying network--at least with respect to network
distances. In addition, will-cut list and distance caches (which store already
probed peers) minimize the unnecessary messages for the network maintenance.
%%
\emph{THANCS} reportedly improves broadcast performance the extent of which, though,
depends on the underlying network since \emph{THANCS} does not provide the
performance guarantee for the improvement of any given overlay topology~\cite{HLY2010}.

@@

While surveying efforts to overcome the mismatch problem
in the area of unstructured \p\ systems,
we came to identify four key methodologies 
utilized by the discussed approaches; they are:
\begin{enumerate}[\itshape i\upshape)]
  \item topology adaptation
  \item forwarding optimization
  \item caching and replication, and
  \item landmarking.
\end{enumerate}
%
A key characteristic shared by many of the presented techniques 
is that they do not strictly adhere to a single methodology.
As approaches attempt to address the challenging problem 
of topology mismatch, they resort to heuristics and so, 
they do not often furnish ``pure'' approaches that exclusively
belong to one of the four aforementioned methodologies.
For instance, \emph{mOverlay} combines topology adaptation 
with landmark binning 
while \emph{Gia} weaves together three methodologies:  topology adaptation, 
forwarding optimization and caching.
%%
The timeline in Figure~\ref{unstructured:timeline} further emphasizes this trend
in the broader picture. All methodologies we discuss,
have been identified and put into practice, early on in the literature,
while in subsequent years, the research community has mostly focused on fine-tuning
the heuristic parameters of these methodologies in pursuit of the
combination that performs best.
This has led to better theoretical 
understanding of the problem and its difficulty. Moreover, the combination of
multiple methodologies has inevitably led to interesting feature/performance trade-offs
that researchers need to tackle in real world applications. For example, 
the effectiveness of a caching/replication component can be undermined
by a continuously adapting overlay that removes important links between peers.
In particular, topology adaptation and forwarding optimization have been constant areas of
focus of the research community. This is of no surprise given 
the volatile nature of unstructured \p systems. 
%%
If a technique is to be evaluated for a specific application domain,
its overall approach as well as its underlying used methodologies have to be considered.
